\section*{Acknowledgements}
This thesis would have been impossible without my supervisor Prof. Sergei Artemov.
He introduced me to the beauty of epistemic logic and its deep relations to proof theory
back at my first years as a graduate student. He inspired and helped shape my topic of research and 
has not stopped being a source of inspiration ever
since. He has provided guidance, support and encouragement through our numerous meetings. 
I want to thank him deeply.

In addition, I want to thank my committee members: Prof. Fitting and Prof. Shankar. Prof. Fitting's classes
 at the Graduate Center have been a tremendous experience and source of knowledge.
The clarity of  both his writings and presentations helped me realize the 
level of academic quality a young student should be striving and aiming toward. Prof. Shankar has
been a great teacher and an inspiring advisor. We spent hours 
discussing topics related to my thesis or to the formal methods area at large. He was one
of the people that invigorated my interest in programming language theory that has
proven very valuable for my research work and my post-graduate career developments.

Prof. Zachos, my external committee member, has been one of the most important
chapters of academic life. It is thanks to him that I attended the PhD program at CUNY
and his guidance in all walks of my academic and non-academic life has contributed 
the most  to its successful completion. He has been an inspiring
scientist, a great mentor and a friend for the last fifteen years that I have 
known him. 


During my studies at the Graduate Center I had the chance to meet 
prominent scientists both from CUNY and worldwide. 
These people helped me expand my scientific horizon and improve
the quality of my work. I would like to thank especially Rohit Parikh, Thomas Wies,  
 Mark Tuttle. I want to give special thanks to Giuseppe Primiero who believed in this work
 since its early stages and co-authored the first related publication with me. 
 The late Kristoffer Rose inspired me with his vigor in programming languages, as well as, his passion
 for open source projects and teaching math and computer science to underrepresented students. 
 In addition, he introduced me to the possibility of high-quality (and fun) research in the industry. 
 He left us early and he is dearly missed.

Moreover, I want to thank my wonderful friends both in NYC and back home. Danai Dragonea, Dafni Anesti, Antonis Bouchagiar 
and Michalis Agathos took great care of me every summer after my teaching obligations
had turned me to a jaded New Yorker. All our beautiful excursions under the
greek sun helped me stay happy and healthy throughout eight intense academic years.
Krystal Raydo and George Rabanca became family  and showed me immense support
through the thick and thin in New York. 
Cagil Tasdemir has been the best friend and a great source of knowledge and wisdom.
I want to thank fellow graduate students (and fellow surfers): Marios Georgiou,
Nikos Melissaris and Ivo Vigan for our great time inside and outside school.
My friend Antonis Stamboulis has been a major support throughout my PhD. 
I had the privilege to discuss with him type theory and modality and having 
him in New York for the last couple of years has been 
the greatest gift. 

Lots of  CUNY staff have worked seamlessly with me to advance my progress.
I want to thank Lina Garcia, and Dilvania Rodriguez. I am deeply indebted to Joe Driscoll 
for his great help and friendship. He accommodated  my schedule to facilitate my teaching which was 
my main means of survival for many years.  I want to thank him dearly.


Finally, I want to thank my kinfolk. Wilson Sherwin's love and understanding has been one of my
most important motivations for completing this work and I am looking forward to our future together.
My family Aliki, Doria, Vaso and Spyros deserve most of the credit for all my achievements including
this work. I had the luck to have two amazing sisters and two loving parents and that's all the wealth 
a man could ask for. 

This dissertation is dedicated to the memory of my late grandparents and grandaunt: 
Aliki Giannioti, a tuberculosis survivor, Theodoros (Lolos) Gianniotis, a fisherman, and
Lambro Mitsiali, a factory  worker, for raising me and my elder sister 
at their home in Anemomylos, Corfu and showing us how to build beauty with as little as
a small house, a fishing boat and a well-kept garden painted fresh white every 
spring.