\documentclass[12pt]{report}
\usepackage[toc,page]{appendix}
\usepackage{lmodern}
\usepackage{dirtytalk}
\usepackage{quotchap}


\usepackage{microtype}

\usepackage{fancyvrb}
\DefineVerbatimEnvironment{code}{Verbatim}{fontsize=\small}
% Package for customizing page layout

% \usepackage{fullpage}

\input{macros}
\usepackage{stmaryrd}
\usepackage{proof-dashed}
\usepackage{tikz}
\usepackage{tikz-cd}
\usepackage[utf8]{inputenc} 
\usepackage{syntax}
\usepackage{amsfonts}
\usepackage{amssymb} 
\usepackage{amsmath}
%\usepackage{amsmath, amsthm, amssymb}
\usepackage{amsthm}
\usepackage{mathpartir}
\usepackage{appendix}
\usepackage{setspace}
\doublespacing
\pagestyle{myheadings}
%opening
\title{Relating Justification Logic Modality and Type Theory in Curry--Howard fashion}
\author{Konstantinos Pouliasis}

\begin{document}

\maketitle


\begin{abstract}
This thesis is a work in the intersection of \ac{JL} and \ac{CHI}. Justification logic is an umbrella of modal logics of knowledge with explicit evidence. 
Justification logics have been used to tackle traditional problems in proof theory (in relation to Godel's provability), philosophy (Gettier examples, Russel's barn paradox). 
The Curry--Howard Isomorphism or \emph{proofs--as--programs} is an understanding of logic that places logical studies in conjunction to type theory and -- in current developments -- 
category theory. The point being that the  understanding a system as a logic, as a typed calculus and as a language of a class of categories constitutes a useful discovery 
that can have many applications.
The application that we will be mainly concerned with is type systems for useful programming language constructions. 
This work  is structured in two parts: 
The first part (CHAPTERS x,y,x) is a revision of my second examination paper and constitutes a bird's eye view into my research topics:  
Logic, Constructive Modality and Type Theory. 
The relevant systems are introduced syntactically together with main  metatheoretic proof techniques which will be useful in the rest of the thesis.
The second part constitues my main contributions: I will propose  a modal type system that extends simple type theory (or, speaking from the logical side of \ac{CHI}, 
intuitionistic propositional logic) with elements of justification logic and will argue about its computational significance. More specifically, I will show  
that the obtained calculus characterizes  certain computational phenomena that abound in modern programming language semantics. 
I will present full metatheoretic
results obtained for this logic/ calculus using and extending techniques from the first part and will provide proofs in the Appendix. In Chapter X, 
I will show the first steps in extending the calculus towards the full propositional universe of justification logic and will provide some obtained metatheoretic results.
Finally, the fourth part exercises an implementation of the calculus as a programming language (to showcase the concept of \emph{proofs--as--programs} in practice). To achieve this ``proof of concept'' result we define our 
language together with its type system
in the metaprogramming framework \say{MAKAM}, which we will discuss briefly, 
and we obtain a type checker for our calculus. Finally, we conclude this work with a small ``outro'' 
where we discuss possibilities for future work and relations of the calculus with other current developments in the theory of functional programming.

This is work is my thesis requirement for my PhD candidacy under the supervision of Distinguished Professor Sergei Artemov at the Department of 
Computer Science of the Graduate Center, CUNY.
\end{abstract}
\include{acknowledge}
\tableofcontents
\include{intro}
\include{Chapter1}
\include{chapter2}
\include{chapter3}
\include{Chapter4}
\include{depjcalc}
\include{makamimpl}
\appendix
\appendixpage
\addappheadtotoc
\include{appendix}





\nocite{Pfenning2009a, Pfenning2009b}


\bibliographystyle{plain}
\bibliography{secondexam}

\end{document}
