\documentclass[letterpaper, 12pt, oneside]{book}
    %\usepackage[toc,page]{appendix}
\usepackage{lmodern}
\usepackage{dirtytalk}
%\usepackage{quotchap}
\usepackage{hyperref}
\usepackage{xparse,nameref}
\usepackage{cleveref}
\usepackage{microtype}

\usepackage{fancyvrb}
\DefineVerbatimEnvironment{code}{Verbatim}{fontsize=\small}
% Package for customizing page layout

% \usepackage{fullpage}

\input{macros}
\usepackage{stmaryrd}
\usepackage{proof-dashed}
\usepackage{tikz}
\usepackage{tikz-cd}
\usepackage[utf8]{inputenc} 
\usepackage{syntax}
\usepackage{amsfonts}
\usepackage{amssymb} 
\usepackage{amsmath}
%\usepackage{amsmath, amsthm, amssymb}
\usepackage{amsthm}
\usepackage{mathpartir}
\usepackage{appendix}
\usepackage{setspace}
\doublespacing
\pagestyle{myheadings}
%opening
\begin{document}

\include{title}
\include{copyright}
\pagenumbering{roman}
\setcounter{page}{3}
\begin{center}
    %A reading of justification logic modality in Curry--Howard Fashion \\
    %by \\
    %Konstantinos Pouliasis
    
    \vspace{.5in}
    This manuscript has been read and accepted by the Graduate Faculty 
    in Computers Science in satisfaction of the dissertation requirement for the degree of Doctor of Philosophy.
    \end{center}
    
    \vspace{1.0in}
    
    \begin{tabular}{p{1.75in}p{0.5in}p{3.5in}}
    \hrulefill                          & &\hrulefill \\
    ~	  			     & & Sergei Artemov\\
    Date                                & & Chair of Examining Committee\\
    ~                                   & & \\
    
    ~                                   & & \\
    \hrulefill                          & &\hrulefill \\
    ~                                   & & Robert Haralick\\
    Date                                & & Executive Officer\\
    \end{tabular}
    
    \vspace{0.75in}
    
    \begin{tabular}{p{1.75in}p{0.5in}p{3.5in}}
    ~ 				& &  Supervisory Committee \\
    ~ 				& &  Melvin Fitting \\
    ~				& &  Subash Shankar \\
    ~				& &  Stathis Zachos \\
    \end{tabular}
    
    
    \vspace{\fill}
    \begin{center}
    \textsc{The City University of New York}
\end{center}
\clearpage

\addcontentsline{toc}{chapter}{Abstract}

\begin{center}
    ABSTRACT \\
    \textsc{Relating Justification Logic Modality and Type Theory in Curry--Howard fashion} \\
    by \\
    \textsc{Konstantinos Pouliasis} \\[0.15in]
    \end{center}
    
    \vspace{0.20in}
    
    \noindent Advisor: Distinguished Professor Sergei Artemov
    
    \vspace{0.20in}
    
    \noindent This thesis is a work in the intersection of \ac{JL} and \ac{CHI}. Justification logic is an umbrella of modal logics of knowledge with explicit evidence. 
    Justification logics have been used to tackle traditional problems in proof theory (in relation to Godel's provability) and philosophy (Gettier examples, Russel's barn paradox). 
    The Curry--Howard Isomorphism or \emph{proofs--as--programs} is an understanding of logic that places logical studies in conjunction with type theory and -- in current developments -- 
    category theory. The point being that  understanding a system as a logic, a typed calculus and,  a language of a class of categories constitutes a useful discovery 
    that can have many applications.
    The applications  we will be mainly concerned with  are type systems 
    for useful programming language constructs. 
    This work  is structured in three parts: 
    The first part (\cref{intui}, \cref{lambda}, \cref{jllogic}) is a  
    a bird's eye view into my 
    research topics:  
    \emph{intuitionistic logic, justified modality} and \emph{type theory}. 
    The relevant systems are introduced syntactically together with main 
     metatheoretic proof techniques which will be useful 
    in the rest of the thesis.
    The second part (\cref{proposal}, \cref{jcalcsem}, \cref{jcalccom}) 
    features my main contributions.
    I will propose  
    a modal type system that extends simple type theory
     (or, isomorphically, intuitionistic propositional logic) with elements of
     justification logic and will argue about its computational significance. 
     More specifically, I will show  
    that the obtained calculus characterizes  certain 
    computational phenomena related to linking (e.g. module mechanisms, foreign function interfaces) 
    that abound in semantics of modern programming languages. 
    I will present full metatheoretic
    results obtained for this logic/ calculus utilizing techniques from the first part 
    and will provide proofs in the Appendix. The Appendix contains also information about
    an implementation of our calculus in the metaprogramming framework {\sf Makam}.
    Finally, I conclude this work with a small ``outro'', ~\cref{ext},
    where I informally show that the ideas underlying my contributions 
    can be extended in interesting ways.
    

%----------------------------------------------------------------------------------------
%	ACKNOWLEDGEMENTS
%----------------------------------------------------------------------------------------

%\setstretch{1.3} % Reset the line-spacing to 1.3 for body text (if it has changed)

\clearpage
{\textcolor{white}{.}   }
\vspace{1.5in}
\thispagestyle{plain}
{\center \em in memory of my grandparents and grandaunt: Aliki, Theodoros and Lambro % Dedication text}
\endcenter} 


\clearpage
\begin{center}
\addcontentsline{toc}{chapter}{Acknowledgements}
{\em \textsc{Acknowledgements}}
\end{center}
\vspace{0.20in}
\include{acknowledge}
\tableofcontents

\addtocontents{toc}{\vspace{2em}}
\mainmatter
\include{intro}
\include{Chapter1}
\include{chapter2}
\include{chapter3}
\include{Chapter4}
\include{order}
\include{Chapter5}
\include{extensions}
\begin{appendices}
\include{appendix}
\end{appendices}




\nocite{Pfenning2009a, Pfenning2009b}


\bibliographystyle{plain}
\bibliography{secondexam}

\end{document}
